---
title: "BASICS OF MACHINE LANGUAGE FINAL ASSIGNMENT"
author: "Harish Kumar uddandi"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(caret)
library(ISLR)
library(factoextra) 
library(tidyverse) 
library(flexclust)
library(cluster)
library(gridExtra)

```


#Loading dataset
```{r}
data<-read.csv("C:/FALL/ML/fuel_receipts_costs_eia923.csv")
str(data)  
```

#selecting attributes 
```{r}

data_df<-data[,c(8,12,13,14,16)]
str(data_df)
colMeans(is.na(data_df))

```

#Data Imputing
```{r}

data_df$fuel_cost_per_mmbtu[is.na(data_df$fuel_cost_per_mmbtu)] <- mean(data_df$fuel_cost_per_mmbtu, na.rm = TRUE) 
colMeans(is.na(data_df))  #all the missing values has been imputed 

```

#.sampling and partition of the data
```{r}
set.seed(2424)
sample_data <- data_df[sample(nrow(data_df), size = 13500, replace = FALSE), ]
train_index <- createDataPartition(sample_data$fuel_cost_per_mmbtu, p=0.75, list = FALSE)
train_data<- sample_data[train_index,]
test_data<- sample_data[-train_index,]



```

# normalization of the data.
```{r}
cluster_data <- train_data %>% select( 'ash_content_pct', 'sulfur_content_pct','fuel_mmbtu_per_unit','fuel_cost_per_mmbtu')

cluster_train <- preProcess(cluster_data, method = "range")
cluster_predict <- predict(cluster_train, cluster_data)

summary(cluster_predict)
```



#Elbow and Silhouette methods are used to find the optimal number of clusters. 
#Elbow Method
```{r}
library(factoextra) # clustering algorithms & visualization
library(flexclust)
fviz_nbclust(cluster_predict,kmeans,method="wss") 
```

#in the plot a clear elbow is at k = 2. Also as the above graph is not clear as it did not show any sharp point at 2. We can use 3 or 4 or 5 as the 'K' value too.

#Silhouttes method
```{r}
#Silhouttes method
fviz_nbclust(cluster_predict,kmeans,method="silhouette")
```

#As observed in elbow method, the optimal clusters identified as 2, but when we have used Silhouttes method, we got the value as 6. As the elbow method was not clear in determining the optimal cluster, we shall use Silhouttes method here
#We have identified the number of clusters. Now we shall apply K-means algorithm

```{r}
#Applying K-means Algorithm
KMean_chk <- kmeans(cluster_predict, centers = 6, nstart = 25) #Number of restarts = 25
```


#Centers
```{r}
KMean_chk$centers

```

#Plotting the cluster using k K-means Algorithm
```{r}
fviz_cluster(KMean_chk, data = cluster_data)
```

#Plotting the Sillohuette average
```{r}
si <- silhouette(KMean_chk$cluster, dist(cluster_predict))
fviz_silhouette(si)
```
#Hence Si(silHouetee coeffient ) value > 0 , i.e 0.83, hence it is a good  clustered.


#The final cluster 
```{r}
fcluster<- KMean_chk$cluster
f_cluster<- cbind(train_data, fcluster)
f_cluster$fcluster<-as.factor(f_cluster$fcluster)
head(f_cluster)

```

# We find the mean of all the quantitative variables 
```{r}
f_cluster%>%group_by(fcluster)%>%
  summarize(
    fuel_mmbtu_per_unit=mean(fuel_mmbtu_per_unit),
            fuel_cost_per_mmbtu=mean(fuel_cost_per_mmbtu),
            sulfur_content=mean(sulfur_content_pct),
    ash_content=mean(ash_content_pct))
```

# Plotting number of ash contents 
```{r}
ggplot(f_cluster) +
  aes(x = fcluster, y = ash_content_pct) +
  geom_jitter(size = 1.2) +
  labs(title = "Ash Content in Clusters") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

```

# Plotting number of clusters 
```{r}
ggplot(f_cluster) +

  aes(x = fcluster, fill = fuel_group_code) +

  geom_bar() +

  scale_fill_brewer(palette = "Accent", direction = 1) +

  labs(x = "Clusters", title = "Classification of Clusters") +

  theme_minimal() +

  theme(

    plot.title = element_text(size = 16L,

    face = "bold",

    hjust = 0.5),

    axis.title.x = element_text(size = 16L,

    face = "bold")

  )
```



#Use multiple-linear regression to determine the best set of variables to predict fuel_cost_per_mmbtu.

#training data 
```{r}
ML_df<- f_cluster
fuel<-ML_df[,-c(1)]
fuel_ML<- preProcess(fuel, method = "range")
fuel_predict <- predict(fuel_ML, fuel)
head(fuel_predict)
```

#performing multiple linear regression model on training data 
```{r}
k<-fuel_predict$fuel_cost_per_mmbtu

 
Z5<- fuel_predict$fuel_mmbtu_per_unit 
Z6<- fuel_predict$sulfur_content_pct
Z7<- fuel_predict$ash_content_pct
```

```{r}
model_check <- lm(fuel_cost_per_mmbtu~.,data=fuel_predict)
summary(model_check)

``` 

#Use the anova analysis
```{r}
anova(model_check)
```


#Test data
```{r}
Check_df<- test_data
fuel<-Check_df[,-c(1)]
fuel_chk<- preProcess(fuel, method = "range")
fuel_check <- predict(fuel_chk, fuel)
head(fuel_check)

```

#performing multiple linear regression model on test data 
```{r}
M<-fuel_check$fuel_cost_per_mmbtu

C6<- fuel_predict$fuel_mmbtu_per_unit 
C7<- fuel_predict$sulfur_content_pct
C8<- fuel_predict$ash_content_pct
```

```{r}
model_check1 <- lm(fuel_cost_per_mmbtu~.,data=fuel_check)
summary(model_check1)

``` 

#Use the anova analysis to predict the model
```{r}
anova(model_check1)
```

The cluster information does not plays an important role to predict fuel_cost_per_mmbtu, since the primary objective of my model  is find the ash content, so cost is not playing crucial role.
